---
title: "Efficacy of a Novel Antihypertensive Drug in treating Stage 1 Hypertension "
author: "Harchit Kaur" 
date: today 
project: 
  output-dir: Output
  execute-dir: project
toc: true 
abstract-title: "Research question:"
abstract: "Does the new drug (AH-27) leads to a greater reduction in SBP?"
format: 
  html: 
    embed-resources: true
    number-sections: true
    code-fold: true
editor: visual 
editor_options: 
  chunk_output_type: console
execute:
  warning: false 
  message: false
  freeze: false
---

## Prepare Session

Load the required libraries, import, screen and correct the data.

```{r}
#| label: prepare session 
#| echo: false 
#| output: false 

  library(pastecs)
  library(ggplot2)
  library(ggpubr); theme_set(theme_bw())
  library(GGally)
  library(performance)
  library(MuMIn)
  library(ggeffects)
  library(tidyverse)
  library(broom)
  library(car)

  library(knitr)                                     # To render tables

# Import data
  BP <- read.csv("data/processed/hyper-tension.csv", stringsAsFactors= T) 

  
str(BP) # shows data typ and preview of the columns
summary(BP) #numeric → mean, min, max, etc.+ factors → counts per category
```

## Explore the data

It's always a good idea to get to know your data, both by exploring them numerically and graphically. Variables in dataset:

-   reduction_sbp (outcome) meassures the reduction of systolic blood pressure in mmHg

-   basline_sbp meassure the baselne systloic blood pressure in mmHg

-   treatment_group shows if the patient received new drug or standard therapy for 8 weeks

### Nummerically

In @tbl-explore-data we see descriptive statistics for the variables of interest.

```{r}
#| label: tbl-explore-data
#| tbl-cap: "Descriptive statistics for the variables of interest"
#| echo: false

sum_stats<-
   stat.desc(BP[, c("baseline_sbp", "reduction_sbp")], norm= T)
 
kable(sum_stats, digits= 1)

```

### Graphically

```{r}
#| label: fig-graphs-1
#| fig-cap: "Overview of relationships within variables"
#| echo: false 

ggpairs(BP, aes(col= treatment_group, alpha= .4))
```

```{r}
#| label: fig-clear-answer
#| fig-cap: "good graph"
#| echo: false 
#this can answer our research question more precisely 
ggarrange(
    ggplot(BP, aes(treatment_group, reduction_sbp)) +
      geom_point(aes(col= treatment_group), size= 3, alpha= .4,
                 position= position_jitter(height= 0, width= .3))+
      geom_boxplot(fill= NA, outliers= F),
    ggplot(BP, aes(treatment_group, baseline_sbp)) +
      geom_point(aes(col= treatment_group), size= 3, alpha= .4) +
      geom_boxplot(fill= NA, outliers= F),
    ncol= 2, nrow= 1, common.legend= T)
```

In @fig-clear-answer we see that both new drug (AH27) and standard care are equally good. But we see that baseline_sbp was slightly higher for standard care. If are hypothesises of higher baseline_sbp leads to higher reduction is true than we can not inteprete that both therapies are equally good. We will have to fit some models and compare them prior concluding something here 😉!

## Fit model

### Formulate and compare

```{r}
#| label:  tbl-fit-models
#| tbl-cap: !expr paste("I choose model 12 becuase it has the lowest AIC", "(AIC=" , AIC))
#| output: false 

#fit models

model.00 <- lm(reduction_sbp ~ 1, data = BP) # ~1 indicates with intercept 
model.01 <- lm(reduction_sbp ~ baseline_sbp, data = BP)
model.02 <- lm(reduction_sbp ~ treatment_group, data = BP)
model.03 <- lm(reduction_sbp ~ age, data = BP)
model.04 <- lm(reduction_sbp ~ sex, data = BP)
model.05 <- lm(reduction_sbp ~ sex + treatment_group, data = BP)
model.06 <-  lm(reduction_sbp ~ age + sex, data = BP)
model.07 <- lm(reduction_sbp ~ treatment_group + age, data = BP)
model.08 <- lm(reduction_sbp ~ baseline_sbp * treatment_group, data = BP)
model.09  <- lm(reduction_sbp ~ sex * treatment_group, data = BP)
model.10 <-  lm(reduction_sbp ~ age * sex, data = BP)
model.11 <- lm(reduction_sbp ~ treatment_group * age, data = BP)
model.12 <- lm(reduction_sbp ~ treatment_group +age +sex + baseline_sbp, data = BP)

kable(compare_performance(model.00, model.01, model.02, model.03, model.04, model.05, model.06,model.07, model.08, model.09, model.10, model.11, model.12 ), digits= 3)

AIC <- AIC(model.12)

# 1088.547 lowest AIC of model.12 
```

### Validate the preferred model

```{r}
#| label: fig-diagnostic-plots
#| fig-cap: "Model diagnostics: these look pretty good "

# First thing: check overdispersion for *lm functions* and glm!!!
  check_model(model.12) 






```

In @fig-diagnostic-plots we see the model represents the reality very well. Homogneity of Variance doesn't look so nice, it points out towards heteroscedasticity meaning that standard errors might be unreliable. The normality of residuals and collinearity look of the model fine though.

## Interpret model

```{r}

```
