---
title: "Efficacy of a Novel Antihypertensive Drug in treating Stage 1 Hypertension "
author: "Harchit Kaur" 
date: today 
project: 
  output-dir: Output
  execute-dir: project
toc: true 
abstract-title: "Research question:"
abstract: "Does the new drug (AH-27) leads to a greater reduction in SBP?"
format: 
  html: 
    embed-resources: true
    number-sections: true
    code-fold: true
editor: visual 
editor_options: 
  chunk_output_type: console
execute:
  warning: false 
  message: false
  freeze: false
---

## Prepare Session

Load the required libraries, import, screen and correct the data.

```{r}
#| label: prepare session 
#| echo: false 
#| output: false 

  library(pastecs)
  library(ggplot2)
  library(ggpubr); theme_set(theme_bw())
  library(GGally)
  library(performance)
  library(MuMIn)
  library(ggeffects)
  library(tidyverse)
  library(broom)
  library(car)

  library(knitr)                                     # To render tables

# Import data
  BP <- read.csv("data/processed/hyper-tension.csv", stringsAsFactors= T) 

  
str(BP) # shows data typ and preview of the columns
summary(BP) #numeric → mean, min, max, etc.+ factors → counts per category
```

## Explore the data

It's always a good idea to get to know your data, both by exploring them numerically and graphically. Variables in dataset:

-   reduction_sbp (outcome) meassures the reduction of systolic blood pressure in mmHg

-   basline_sbp meassure the baselne systloic blood pressure in mmHg

-   treatment_group shows if the patient received new drug or standard therapy for 8 weeks

### Nummerically

In @tbl-explore-data we see descriptive statistics for the variables of interest.

```{r}
#| label: tbl-explore-data
#| tbl-cap: "Descriptive statistics for the variables of interest"
#| echo: false

sum_stats<-
   stat.desc(BP[, c("baseline_sbp", "reduction_sbp")], norm= T)
 
kable(sum_stats, digits= 1)

```

### Graphically

```{r}
#| label: fig-graphs-1
#| fig-cap: "Overview of relationships within variables"
#| echo: false 

ggpairs(BP, aes(col= treatment_group, alpha= .4))
```

```{r}
#| label: fig-clear-answer
#| fig-cap: "good graph"
#| echo: false 
#this can answer our research question more precisely 
ggarrange(
    ggplot(BP, aes(treatment_group, reduction_sbp)) +
      geom_point(aes(col= treatment_group), size= 3, alpha= .4,
                 position= position_jitter(height= 0, width= .3))+
      geom_boxplot(fill= NA, outliers= F),
    ggplot(BP, aes(treatment_group, baseline_sbp)) +
      geom_point(aes(col= treatment_group), size= 3, alpha= .4) +
      geom_boxplot(fill= NA, outliers= F),
    ncol= 2, nrow= 1, common.legend= T)
```

In @fig-clear-answer we see that both new drug (AH27) and standard care are equally good in reduction of systolic blood pressure. But we see that baseline_sbp was slightly higher for standard care. If are hypothesises of higher baseline_sbp leads to higher reduction is true than we can not inteprete that both therapies are equally good. We will have to fit some models and compare them prior concluding something here !

## Fit model

### Formulate and compare

```{r}
#| label:  fit-models
#| output: false 

#fit models

model.00 <- lm(reduction_sbp ~ 1, data = BP) # ~1 indicates with intercept 
model.01 <- lm(reduction_sbp ~ baseline_sbp, data = BP)
model.02 <- lm(reduction_sbp ~ treatment_group, data = BP)
model.03 <- lm(reduction_sbp ~ age, data = BP)
model.04 <- lm(reduction_sbp ~ sex, data = BP)
model.05 <- lm(reduction_sbp ~ sex + treatment_group, data = BP)
model.06 <-  lm(reduction_sbp ~ age + sex, data = BP)
model.07 <- lm(reduction_sbp ~ treatment_group + age, data = BP)
model.08 <- lm(reduction_sbp ~ baseline_sbp * treatment_group, data = BP)
model.09  <- lm(reduction_sbp ~ sex * treatment_group, data = BP)
model.10 <-  lm(reduction_sbp ~ age * sex, data = BP)
model.11 <- lm(reduction_sbp ~ treatment_group * age, data = BP)
model.12 <- lm(reduction_sbp ~ treatment_group + age + sex + baseline_sbp, data = BP)

kable(compare_performance(model.00, model.01, model.02, model.03, model.04, model.05, model.06,model.07, model.08, model.09, model.10, model.11, model.12 ), digits= 3)

AIC <- AIC(model.12) 
# 1088.547 lowest AIC of model.12 
```

I choose model.12 because it has the lowest AIC= 1088.55 , which tells us that its the best model to work with for interpretating results.

### Validate the preferred model

```{r}
#| label: fig-diagnostic-plots
#| fig-cap: "Model diagnostics: these look pretty good "

# First thing: check overdispersion for *lm functions* and glm!!!
  check_model(model.12) 

```

In @fig-diagnostic-plots we see the model represents the reality very well. Homogneity of Variance doesn't look so nice, it points out towards heteroscedasticity meaning that standard errors might be unreliable. The normality of residuals and collinearity look of the model fine though.

## Interpret model

```{r}
#| label: fig-interpret-objects
#| fig-cap: "effectiveness of the treatments"
#| echo: false 


## Coefficients, SE, z and p -----
  coef.table<- round(summary(model.12)$coefficients, 3)

## Approximate R2 ----- amount of variablity explained by your model
  r.2<- round(attributes(r.squaredLR(model.12))$adj.r.squared, 3)  

## Compare performance to null-model ----- stored the anova of glm.00 and gln.02 in null.chi - - in quarto you extract number from model and store them in objects/names like null.chi
  null.chi<- anova(model.12, model.00)

# Objects to plot predictions -----------------------------------------------

## Interaction plot -----
  # Create vector of values for x-axis (to get smooth lines) 
  t.vec<- seq(min(BP$baseline_sbp), max(BP$baseline_sbp), length.out= 100)
  
  # Calculate min, mean max for colour coding of spice at min, mean, max
  #summary(BP$age)       # The min, mean and max need to be entered by
                            # hand in the predict_response() function below!!!!
  
  # Create predictions dataframe
  pred.int<- predict_response(model.12,
                              c("baseline_sbp [t.vec]",
                                "age [43, 55, 67]"))
  # Assign text labels to the th groups
  levels(pred.int$group)<- c("young", "midage", "old")
  
  # Create plot - created a plot which is stored under the name pred.plot 
 pred.plot <- ggplot(pred.int, aes(x, predicted)) +
  
  # actual data
  geom_point(data = BP,
             aes(baseline_sbp, reduction_sbp, col = treatment_group),
             alpha = 0.4) +
  
  # confidence bands
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high,
                  fill = group),
              alpha = 0.2) +
  
  # prediction lines
  geom_line(aes(col = group), linewidth = 1) +
  
  labs(
    x = "baseline systolic\n blood pressure (mmHg)",
    y = "reduction of systolic\n blood pressure (mmHg)",
    col = "Age",
    fill = "Age"
  ) +
  
  theme_minimal()

```

```{r}
#| label: tbl-glm 
#| tbl-cap:  !expr paste("Estimates, associated standard errors and statistical significance of the parameters estimated by a LM.", "($R^2$=", r.2, ", $\\chi^2_{(", null.chi[2, 3], ",", null.chi[2, 1], ")}$=", round(null.chi[2,4], 3), ", $p$", format.pval(null.chi[2, 5], eps = 0.001), ")") 
#| echo: false 

# Create table
kable(coef.table)
```

A Linear Model with all predictors of the data (*R*^2^= `r r.2` , `r paste("$\\chi^2_{(", null.chi[2,3], ",", null.chi[2,1], ")}$")`= `r round(null.chi[2, 4], 3)` , *p*`r format.pval(null.chi[2, 5], eps= 0.001)`; @tbl-glm) performed significantly better than a null-model, comprising an intercept only. We can see that there is a small sex response difference but it is not significant: $\beta_{age}$= `r round(coef.table[4, 1], 3)` , *SE*= `r round(coef.table[4, 2], 3)` , *z*= `r round(coef.table[4, 3], 3)`, *p*= `r round(coef.table[4,  4], 3)`.

This findings is corroborated by @fig-predict, in which we see that, that older patients have reduced reponse to treatment irrespective of the drug type. We also see that the new drug is slightly better than the standardcare therapy.

```{r}
#| label: fig-predict
#| fig-cap: "Raw data and model predictions to visualize the relationship baseline systelic blood pressure and recution of it after giving treatment depending on age and sex"
#| echo: false 

# Plot predictions
  pred.plot
#summary(model.12)
```

```{r}
#| label: store session
#| echo: false 
#| output: false 

#do it in the end 
version<- sessionInfo() #for reproducibility
```
